Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-27 TACAS2018/paper.tex
sec 6

Stability denotes the property of a model to remain within a bounded set of states around an equilibrium point. 
In this paper, we are interested in asymptotic stability, where the output is expected to converge to a certain steady state value.  
We specify model stability via Jury's criterion~\cite{fadali}\footnote{More details on Jury's criterion is in Appendix~\ref{sec:closed_stability}}, 
which allows for an expression with an algebraic formula. 
We transform the model into a representation that contains the coefficients of the characteristic polynomial (used by Jury's criterion) expressed as sums of controller and plant values (with no multiplications, which are expensive for the SAT solver).  
Such a representation is called Controllable and Observable Canonical Form~\cite{Astrom08} \footnote{See parts~\ref{sec:reachable} and~\ref{sec:observable} in the Appendix}.  
In this Section we explore the implications of using these canonical forms alongside an FWL representation.

We assume that the domain of the observer model is further discretised, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},{\mat{C}_d}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, where $*$ represents the appropriate
dimension for each matrix.
New problems arise due to the FWL effects on these matrices.  The Canonical
form of the resulting system is expressed by the formula:
%
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\left [\begin{array}{c} \vec{r}_k + \vec{\eta}_2\\ \vec{\eta}_1\end{array}\right]\nonumber\\
\vec{y}_k=&\mat{C}_d\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
where $\mat{T}_c$ and $\mat{T}_o$ trasform the system into Canonical Form~\cite{Astrom08}, $\vec{x}'$ and $\vec{x}_e'$ are the state of the system and the error of the observer evaluated in the transformed domain \note{this statement will not be understood} and
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle}\right)\mat{T}_c^{-1}&
\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
\mat{C}_{e}=&\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}
\end{align*}
}
{\setlength{\abovedisplayskip}{0pt}
\begin{align*}
\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\\
\eta_1=&\nu_1+\nu_1'+\nu_t& \eta_2=&\nu_2+\nu_2',
\end{align*}
}
\note{it is very hard if not impossible to understand these operations without having thoroughly gone through the whole Appendix, which is unlikely for a reviewer to do.}
where $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations (see Section~\ref{sec:quantization-noise}).

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Safety of a digitally observed system}\label{sec:cof_safety}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
A safety specification gives raise to a requirement on the
states of the model, such that the controller
must ensure that the state never violates the
requirement.
In~this work, the safety property is expressed as:
%
\begin{align}
  \phi_\mathit{safety}: = & \underline{y} \leq y_k \leq \overline{y}
\equiv \forall k\geq 0.\, \mat{G}\vec{x}_k \leq \vec{h},
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
Note that FWL effects are not explicitly captured here
as they are implicit from Equation \ref{eq:observer_LTI_FWL_cf}.
%% We assume that we do not want saturation in our
%% system, hence the control loop has only one mode.
%% First we must
%% translate the safety specification into the statespace.  Replacing on the
%% second part of equation \eqref{eq:observer_LTI_cf} we obtain
%% %
%% \begin{equation}
%% \phi_{safety} := \underline{y}<\mat{C}\vec{x}_k<\overline{y}, \forall k \geq 0
%% \label{eq:of_bounds}
%% \end{equation}

Furthermore, it is practically relevant to consider the 
constraints $\phi_\mathit{input}$ on the input
signal $u_{k}$ and $\phi_\mathit{init}$ on the initial states $x_0$,
which we assume have given bounds, namely 
$\phi_\mathit{input} := \underline{u} \leq u_k \leq \overline{u}$, 
which comes from physical restrictions in the energy of the system; and
$\phi_\mathit{init}\equiv \mat{G}' \vec{x}_0 \leq \vec{h}'$,
where $\mat{G}'$ and $\vec{h}'$ describe the initial state-space.

>>>>

Stability denotes the property of a model to unfold staying within a bounded set around an equilibrium point. 
In this paper, we are interested in asymptotic stability, where the model is further expected to converge to the steady state value (i.e., the equilibrium point).   
%
We specify model stability via Jury's criterion~\cite{fadali}\footnote{More details on Jury's criterion is in Appendix~\ref{sec:closed_stability}.}, 
which boils down to a requirement expressed as an algebraic formula - a SAT solver will manipulate its coefficients. 
%
We transform the model into a representation that contains the coefficients of its characteristic polynomial, which are used in Jury's criterion, 
that are expressed as sums of controller and plant values: 
this representation has no multiplications, which are expensive for a SAT solver.   
%
Such a representation is known as 
% is called Controllable and Observable 
``canonical form''~\cite{Astrom08}\footnote{See parts~\ref{sec:reachable} and~\ref{sec:observable} in the Appendix.} 
and is dynamically equivalent to the given model.  
In this Section we explore the implications of using this canonical form, 
alongside the use of a FWL representation.

\medskip

We assume that the domain of the observer model is further discretised, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},{\mat{C}_d}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, where $*$ represents the appropriate
dimension for each matrix.
New problems arise due to the FWL effects on these matrices.  
The canonical form of the resulting model is expressed by:
%
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\left [\begin{array}{c} \vec{r}_k + \vec{\eta}_2\\ \vec{\eta}_1\end{array}\right]\nonumber\\
\vec{y}_k= &\mat{C}_d\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
where the matrices $\mat{T}_c$ and $\mat{T}_o$ linearly transform the model into canonical form~\cite{Astrom08}, $\vec{x}'$ and $\vec{x}_e'$ are the state of the system and the error of the observer evaluated in the transformed domain, 
and we have introduced the following terms: 
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle}\right)\mat{T}_c^{-1}&
\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
\mat{C}_{e}=&\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}
\end{align*}
}
{\setlength{\abovedisplayskip}{0pt}
\begin{align*}
\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\\
\eta_1=&\nu_1+\nu_1'+\nu_t& \eta_2=&\nu_2+\nu_2',
\end{align*}
}
\note{it is very hard if not impossible to understand these operations without having thoroughly gone through the whole Appendix, which is unlikely for a reviewer to do.}
Here $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations (see Section~\ref{sec:quantization-noise}).

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Safety of a digitally observed system}\label{sec:cof_safety}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
A safety specification gives raise to a requirement on the
states of the model, such that the controller
must ensure that the state never violates the
requirement.
In~this work, the safety property is expressed as:
%
\begin{align}
  \phi_\mathit{safety}: = & \, \underline{y} \leq y_k \leq \overline{y}
\equiv \forall k\geq 0 \, \mat{G}\vec{x}_k \leq \vec{h},
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
Note that FWL effects are not explicitly captured here
as they are implicit from Equation \ref{eq:observer_LTI_FWL_cf}.
%% We assume that we do not want saturation in our
%% system, hence the control loop has only one mode.
%% First we must
%% translate the safety specification into the statespace.  Replacing on the
%% second part of equation \eqref{eq:observer_LTI_cf} we obtain
%% %
%% \begin{equation}
%% \phi_{safety} := \underline{y}<\mat{C}\vec{x}_k<\overline{y}, \forall k \geq 0
%% \label{eq:of_bounds}
%% \end{equation}

Furthermore, it is practically relevant to consider the 
constraints $\phi_\mathit{input}$ on the input
signal $u_{k}$ and $\phi_\mathit{init}$ on the initial states $x_0$,
which we assume have given bounds, namely 
$\phi_\mathit{input} := \underline{u} \leq u_k \leq \overline{u}$, 
which comes from physical restrictions in the energy of the system; and
$\phi_\mathit{init}\equiv \mat{G}' \vec{x}_0 \leq \vec{h}'$,
where $\mat{G}'$ and $\vec{h}'$ limits the initial conditions of the model.
